---
title: "Week 1: Introduction to Statistics and R"
subtitle: "PLS 120 - Applied Statistics in Agriculture"
output: html_notebook
---

# Welcome to R! 🌾

This is your first R Markdown notebook. You can run code chunks by clicking the green play button or pressing Ctrl+Shift+Enter.

## R as a Calculator

Let's start with basic arithmetic operations:

```{r basic_math}
# Addition
2 + 3

# Subtraction
10 - 4

# Multiplication
5 * 6

# Division
20 / 4

# Exponentiation
2^3

# Square root
sqrt(16)
```

## Creating Variables

In R, we can store values in variables using the assignment operator `<-`:

```{r variables}
# Store crop yield data
wheat_yield <- 4.2
corn_yield <- 8.5
soybean_yield <- 3.1

# Display the values
wheat_yield
corn_yield
soybean_yield

# Calculate total yield
total_yield <- wheat_yield + corn_yield + soybean_yield
total_yield
```

## Working with Vectors

Vectors are collections of values. Very useful for agricultural data!

```{r vectors}
# Create a vector of plant heights (in cm)
plant_heights <- c(25, 30, 28, 32, 27, 29, 31, 26)
plant_heights

# Calculate basic statistics
mean(plant_heights)      # Average height
max(plant_heights)       # Tallest plant
min(plant_heights)       # Shortest plant
length(plant_heights)    # Number of plants
```

## Fun Agricultural Calculator

Let's create some useful agricultural calculations:

```{r ag_calculator}
# Farm area calculation
field_length <- 100  # meters
field_width <- 50    # meters
field_area <- field_length * field_width
cat("Field area:", field_area, "square meters\n")

# Convert to hectares (1 hectare = 10,000 m²)
field_hectares <- field_area / 10000
cat("Field area:", field_hectares, "hectares\n")

# Seed calculation
seeds_per_sqm <- 200
total_seeds_needed <- field_area * seeds_per_sqm
cat("Total seeds needed:", total_seeds_needed, "seeds\n")

# Convert to kg (assuming 1000 seeds = 50g)
seed_weight_kg <- (total_seeds_needed / 1000) * 0.05
cat("Seed weight needed:", seed_weight_kg, "kg\n")
```

## Your Turn! 🚜

Try these exercises:

```{r exercise1}
# Exercise 1: Calculate fertilizer needed
# If you need 150 kg of fertilizer per hectare, 
# how much do you need for a 2.5 hectare field?

hectares <- 2.5
fertilizer_per_hectare <- 150
total_fertilizer <- # Complete this calculation

# Uncomment the line below when you're done:
# cat("Total fertilizer needed:", total_fertilizer, "kg\n")
```

```{r exercise2}
# Exercise 2: Crop yield comparison
# Create vectors for yields of three varieties
variety_A <- c(4.2, 4.5, 4.1, 4.3, 4.4)
variety_B <- c(3.8, 4.0, 3.9, 4.1, 3.7)
variety_C <- c(4.6, 4.8, 4.4, 4.7, 4.5)

# Calculate the average yield for each variety
avg_A <- # Complete this
avg_B <- # Complete this  
avg_C <- # Complete this

# Which variety has the highest average yield?
```

## Loading Data

Let's load some real agricultural data:

```{r load_data}
# Load the sample crop data
crop_data <- read.csv("sample_crop_data.csv")

# Display first few rows
head(crop_data)

# Get basic information about the dataset
str(crop_data)
summary(crop_data)
```

## Simple Visualization

```{r simple_plot}
# Create a simple histogram of yields
hist(crop_data$yield, 
     main = "Distribution of Crop Yields",
     xlab = "Yield (tons/ha)",
     col = "lightgreen",
     border = "darkgreen")
```

## Next Steps

Great job! You've learned:
- ✅ Basic R arithmetic
- ✅ Creating variables and vectors
- ✅ Agricultural calculations
- ✅ Loading and exploring data
- ✅ Creating simple plots

**Next week:** We'll dive deeper into descriptive statistics and data visualization!

---

*Click the "Knit" button above to generate a nice HTML report of your work!*